---
title: "Creating your own R package"
author: "Martin Law"
date: "2026-09-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, eval=FALSE}
rmarkdown::render("create_R_package_guide.Rmd", output_format = "word_document")
```


# Prerequisites

Assumption: you use github in your day-to-day workflow. Easiest approach:

 - Create repo on github
 - New project using this github repo
 
See [Happy git with R](https://happygitwithr.com/new-github-first.html) for more details

# Turn existing project into a package

```{r, eval=FALSE}
install.packages("librarian")
librarian::shelf(devtools, usethis, roxygen2, here)
```

 - `usethis::create_package(here())` -- turns project into a package
 - Copy any existing functions into `R/`
 - `devtools::load_all()` -- attaches your package (like `library`) 
 - Make edits to your functions, like specifying imported packages within code
 - `devtools::use_mit_license()` -- add MIT licence document (other licences are available)
 - Edit DESCRIPTION file -- your name, email address, etc.
 - `usethis::use_package("otherpkg")` if your your package uses another package 
 - Insert Roxygen skeleton (Alt+Ctrl+Shift+R) -- creates templates for help pages
 - Fill out templates for help pages
 - `devtools::document()` or `roxygen2::roxygenise()` -- creates the actual help page files
 - Optional: `usethis::use_readme_rmd()` -- creates a template README R  Markdown file
 - Optional: `usethis::build_readme()` -- renders the R Markdown file, creating the README.md file
 - `devtools::check()` -- builds the package and checks for problems
 - `devtools::install()` -- installs the package proper
 

# Make the package available via github

 - Commit
 - Push
 - Users can then install by running `remotes::install_github("username/pkgname")`


# Update your package in future

 - `devtools::load_all()` -- attaches your package as before
 - Make edits
 - `devtools::check()` 
 - Commit changes
 - `usethis::use_version()` -- update the version number
 - Push
 - On github, create a new Release, using the new version number
 
# Next steps

 - Create tests, to verify your functions work as intended
 - Create vignettes, to explain how your functions work in more detail
